---
title: "United States Solar Energy Consumption"
output: html_document
---
# Introduction
This script explores trends in monthly United States solar energy consumption. Data was collected by the U.S. Energy Information Administration (https://www.eia.gov/).
```{r setup, include=FALSE}
#Install packages
library(httr)
library(jsonlite)
library(dplyr)
library(Metrics)
library(zoo)
library(knitr)
library(kableExtra)
library(scales)
library(forecast)
```

# Setup
Define Functions:
```{r function-definitions}
#Function to convert a PERIOD_ID (YYYYMM) to a date (MM/01/YYYY)
period_to_date <- function(col){
  DATE <- as.Date(paste(as.numeric(col) %% 100, 1,
                                  round(as.numeric(col) %/% 100,0),
                                  sep = "/"), "%m/%d/%Y")
  return(DATE)
}

#Function to pull data using API call
API_request <- function(API_call, API_key){
  res = GET(sub('YOUR_API_KEY_HERE', API_key, API_call))
  data = fromJSON(rawToChar(res$content))
  consumption_data <- as.data.frame(data$series$data)
  consumption_data <- consumption_data %>% 
    rename(PERIOD_ID = X1,
      TRILLION_BTU = X2)
  consumption_data$DATE <- period_to_date(consumption_data$PERIOD_ID)
  return(consumption_data)
}

#Function to run naive model
naive_model <- function(ts_train, ts_test){
  require(forecast)
  model <- naive(ts_train, level=c(80,95), h=12)
  plot(model)
  lines(ts_test, col='red')
  data_eval <- data.frame(actual_data=as.matrix(ts_test),
                          date=as.Date(as.yearmon(time(ts_test))))
  data_eval <- cbind(data_eval, predicted = model$mean)
  return(list(aicc = model$aicc, model = model, data_eval = data_eval))
}

#Function to run ARIMA model without external regressors
arima_model <- function(ts_train, ts_test){
  require(forecast)
  model <- auto.arima(ts_train,
                          stepwise = FALSE,
                          approximation = FALSE,
                          seasonal = FALSE,
                          ic = 'aicc')
  forecast_data <- forecast(model, level=c(80,95), h= 12)
  plot(forecast_data)
  lines(ts_test, col='red')
  data_eval <- data.frame(actual_data=as.matrix(ts_test),
                          date=as.Date(as.yearmon(time(ts_test))))
  data_eval <- cbind(data_eval, predicted = forecast_data$mean)
  return(list(aicc = model$aicc, model = model, data_eval = data_eval))
}

#Function to run ARIMA model with Fourier component of different external regressors and pull out best performing forecast by aicc
best_fourier <- function(ts_train, ts_test){
  require(forecast)
  models <- lapply(seq(1,6), function(j){
    try(auto.arima(ts_train,
                   xreg = fourier(ts_train, K=j),
                   seasonal = FALSE,
                   stepwise = FALSE,
                   approximation = FALSE,
                   ic = 'aicc'))
  })
  best_forecast <- which.min(sapply(models, function(x){x[['aicc']]}))
  forecast_data <- forecast(models[[best_forecast]], level=c(80,95),
                                xreg = fourier(ts_train, K=best_forecast,
                                               h= 12))
  plot(forecast_data)
  lines(ts_test, col='red')
  data_eval <- data.frame(actual_data=as.matrix(ts_test),
                          date=as.Date(as.yearmon(time(ts_test))))
  data_eval <- cbind(data_eval, predicted = forecast_data$mean)
  return(list(aicc = models[[best_forecast]]$aicc,
              model = models[[best_forecast]], k= best_forecast,
              data_eval = data_eval))
}

# Function to run ETS model
ets_model <- function(ts_train, ts_test){
  require(forecast)
  model <- ets(ts_train)
  forecast_data <- forecast(model, level=c(80,95), h= 12)
  plot(forecast_data)
  lines(ts_test, col='red')
  data_eval <- data.frame(actual_data=as.matrix(ts_test),
                          date=as.Date(as.yearmon(time(ts_test))))
  data_eval <- cbind(data_eval, predicted = forecast_data$mean)
  return(list(aicc = model$aicc, model = model, data_eval = data_eval))
}

#Function to get model metrics
get_metrics <- function(data_eval){
  model_MAPE <- mape(data_eval$actual, data_eval$predicted)
  model_RMSE <- round(rmse(data_eval$actual, data_eval$predicted), 2)
  model_MAE <- round(mae(data_eval$actual, data_eval$predicted), 2)
  model_MASE <- round(mase(data_eval$actual, data_eval$predicted), 2)
  model_bias <- round(bias(data_eval$actual, data_eval$predicted),2)
  metrics <- list(model_MAPE, model_RMSE, model_MASE, model_MAE, model_bias)
  names(metrics) <- c('MAPE', 'RMSE', 'MAE', 'MASE', 'Bias')
  metrics <- as.data.frame(metrics)
  metrics$MAPE <- percent(metrics$MAPE, accuracy = 0.01)
  return(metrics)
}

#Create summary table of all metrics results
metrics_summary <- function(dfNames) {
  do.call(rbind, lapply(dfNames, function(x) {
    cbind(Model = strsplit(x,'_')[[1]][1], get(x))
  }))
}
```

Pull the data using the API request.
```{r data-pull}
api_key = 'b7369a57166eee9b3612e8774f172306'
solar <- API_request("http://api.eia.gov/series/?api_key=YOUR_API_KEY_HERE&series_id=TOTAL.SOTCBUS.M", api_key)
```

# Data Exploration
Convert consumption data to numeric, impute missing data with 0, and slice the data to remove time points before the first occurrence of solar energy consumption (where TRILLION_BTU > 0).
```{r data-cleaning}
solar$TRILLION_BTU <- as.numeric(solar$TRILLION_BTU)
solar$TRILLION_BTU[is.na(solar$TRILLION_BTU)] <- 0
solar <- solar[order(solar$DATE),]
first_value <- which.max(solar$TRILLION_BTU > 0)
solar <- solar %>% slice(first_value:n())
```
View the time series plot for solar energy consumption.
```{r plot-full-data}
plot(solar$DATE, solar$TRILLION_BTU, type = 'l')
```

Split data into train/test sets - leaving 1 year of data for performance evaluation.
```{r train-test-split}
#split where there are 12 remaining months left
cut_loc <- which(solar$DATE > seq(max(solar$DATE), length=2, by="-1 years")[2])
train_data <- solar[-cut_loc,]
test_data <- solar[cut_loc,]
```
Create time series data train/test sets.
```{r time series creation}
start_year_train = as.numeric(format(as.Date(min(train_data$DATE), format="%d/%m/%Y"),"%Y"))
start_month_train = as.numeric(format(as.Date(min(train_data$DATE), format="%d/%m/%Y"),"%m"))
ts_train <- ts(train_data$TRILLION_BTU,
               start = c(start_year_train, start_month_train), frequency = 12)

start_year_test = as.numeric(format(as.Date(min(test_data$DATE), format="%d/%m/%Y"),"%Y"))
start_month_test = as.numeric(format(as.Date(min(test_data$DATE), format="%d/%m/%Y"),"%m"))
ts_test <- ts(test_data$TRILLION_BTU,start = c(start_year_test, start_month_test), frequency=12)
```

Explore the components of the time series train data.
```{r ts-decomposition}
fit <- stl(ts_train, s.window="period")
plot(fit)
```
In the plots above, there is a similar seasonal trend throughout the entire time series. The trend in solar energy consumption begins increasing slowly in the late 1980's and then more rapidly increases beginning around 2012. The more rapid increase in 2012 may be due to policy changes in the United States including Production Tax Credits and Investment Tax Credits for solar energy and the American Reinvestment and Recovery Act of 2009.
```{r}
# additional plots
seasonplot(ts_train)
```
The seasonal plot above shows that the most solar energy is consumed during the summer. This trend is most likely due to the availability and production of solar power during the sunnier time of year.
# Forecast Models
Now that I have explored the trends in the data, I will begin exploring forecast models.
First, I will create a baseline forecast using a naive model.
```{r}
naive_results <- naive_model(ts_train, ts_test)
```
In addition to viewing the forecast plot, I will also quantify the model performance using several model metrics, including mean absolute percentage error (MAPE), root mean squared error (RMSE), mean absolute error (MAE), mean absolute scaled error (MASE) and bias. 
```{r}
naive_metrics <- get_metrics(naive_results$data_eval)
kable(as.data.frame(naive_metrics) %>% select_all, caption = "Naive Forecast Metrics") %>% kable_styling(full_width = F, bootstrap_options = c("condensed"))
```
Now that the baseline is established, I will experiment with different time series forecasting models to find the best performing method. I will begin with an ARIMA model:
```{r}
best_arima_result <- arima_model(ts_train, ts_test)
arima_metrics <- get_metrics(best_arima_result$data_eval)
kable(as.data.frame(arima_metrics) %>% select_all, caption = "ARIMA Forecast Metrics") %>% kable_styling(full_width = F, bootstrap_options = c("condensed"))
```

Next, I will try adding Fourier components to the ARIMA forecast:
```{r}
best_fourier_result <- best_fourier(ts_train, ts_test)
fourier_metrics <- get_metrics(best_fourier_result$data_eval)
kable(as.data.frame(fourier_metrics) %>% select_all, caption = "ARIMA/Fourier Forecast Metrics") %>% kable_styling(full_width = F, bootstrap_options = c("condensed"))
```

Now, I will generate an ETS model:
```{r}
best_ets_result <- ets_model(ts_train, ts_test)
ets_metrics <- get_metrics(best_ets_result$data_eval)
kable(as.data.frame(ets_metrics) %>% select_all, caption = "ETS Forecast Metrics") %>% kable_styling(full_width = F, bootstrap_options = c("condensed"))
```
Print out the summary statistics below to compare the forecast models:
```{r}
summary_stats <- metrics_summary(c('naive_metrics', 'arima_metrics', 'fourier_metrics', 'ets_metrics'))
kable(as.data.frame(summary_stats) %>% select_all, caption = "Summary Forecast Metrics") %>% kable_styling(full_width = F, bootstrap_options = c("condensed", "striped")) %>%
  row_spec(which(summary_stats$RMSE == min(summary_stats$RMSE)),
           bold = T, color = "white", background = "green") %>%
  column_spec(1, bold=T)
```
The table above shows that the ETS model had the best performance for this data set; the MAPE, RMSE, MAE, and MASE were all lowest with this model. The ARIMA model had the least bias, but higher values for the other metrics. The models are over-predicting the energy consumption, as can be seen from the positive bias values. 
# Final Forecast
I can now apply the ETS model to predict United States solar energy consumption for the next 10 years.
```{r}
forecast_data <- forecast(best_ets_result$model, level=c(80,95), h= 120)
plot(forecast_data, xlab="Trillion MMBtu", ylab="y-axix label")
```